---
title: "Create/enhance data visualisations with scRUtils"
author: "I-Hsuan Lin"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
package: scRUtils
output:
  BiocStyle::html_document:
    toc_float: yes
vignette: >
  %\VignetteIndexEntry{2. General data visualisations}
  %\VignetteEngine{knitr::rmarkdown}
  %VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(error = FALSE, warning = FALSE, message = TRUE, tidy = FALSE,
                      cache = FALSE, dpi = 100)
library(BiocStyle)
```

# Overview

`scRUtils` provides various utilities for visualising and functional analysis of RNA-seq data,
particularly single-cell dataset. It evolved from a collection of helper functions that were
used in our in-house scRNA-seq processing workflow.

The documentation of this package is divided into 5 sections:

1. [Introduction to scRUtils](https://ycl6.github.io/scRUtils/scRUtils-intro.html)
2. [General data visualisations](https://ycl6.github.io/scRUtils/basics.html)
3. [Single-cell visualisations](https://ycl6.github.io/scRUtils/singleCell.html)
4. [Markers and DEGs](https://ycl6.github.io/scRUtils/markers.html)
5. [Demo datasets](https://ycl6.github.io/scRUtils/demoData.html)

This vignette (#2) will demonstrate functions for general visualisation purposes.

# Load packages

To use `scRUtils` and relevant packages in a R session, we load them using the `library()` command.

```{r load-packages, message = FALSE}
library(scRUtils)

library(ggforce)
library(ggplot2)
library(scater)
```

# Usage

## Jupyter Notebook-specific

### Change `repr.plot.*` behaviour

The `fig()`, function uses `options()` to change the behaviour of `repr.plot.*`. It provides a
quick and easy way to change plot size and other `repr.plot.*` behaviours when running R in a
**Jupyter Notebook**. If use without indicating any argument, the plot behaviour will be reset
to default. The `reset.fig()` is an alias of `fig()`.

The example below is not evaluated as the function has no effect in R Markdown.

```{r repr-plot, eval = FALSE}
library(ggplot2)

# Change plot area width to 8 inches and height to 5 inches
fig(width = 8, height = 5)
ggplot(mpg, aes(class)) + geom_bar()

# Reset to default settings
fig()

# Change plot wider and taller
fig(width = 14, height = 10)
ggplot(mpg, aes(class)) + geom_bar()

# Alias of fig()
reset.fig()
```

## Discrete palette

### The `c30()` and `c40()` palettes

The `c30()` palette has 30 unique colours and `c40()` palette has 40 unique colours.
The `c40()` colour palette is taken from `plotScoreHeatmap()` of the `r Biocpkg("SingleR")`
package (which itself is based on \pkg{DittoSeq} and Okabe-Ito colors).

```{r show-palettes, fig.small = TRUE}
# Show colours as pie charts
pie(rep(1,30), col = c30(), radius = 1.05)
pie(rep(1,40), col = c40(), radius = 1.05)
```

### Choose discrete colours

The `choosePalette()` function takes a character vector of features and optionally a vector of
color codes to evaluate if the supplied color codes has sufficient number of colours. It returns
a named vector of color codes based on the input features, with the same length as the unique
features.

By default, it uses the `c30()` palette when no more than 30 colours are required, then the
`c40()` palette, and lastly the `rainbow()` colour palette when requiring more than 40 colours.

The example below shows using a character vector of 10 letters as input and `choosePalette()`
returns 5 colours.

```{r choose-palette-vector}
feat <- rep(LETTERS[1:5], 2)
feat
choosePalette(feat) # use c30()
```

Next example shows using a factor of 15 letters and 3 levels as input and `choosePalette()`
returns 3 of the 10 colours from the `rainbow(10)` colour palette.

```{r choose-palette-factor}
feat <- factor(rep(LETTERS[1:3], 5))
feat
choosePalette(feat, rainbow(10))
```

## Parallel sets diagram

### Add nudged labels in a parallel sets diagram

The `geom_parallel_sets_labs()` function in this package is the same function as
`geom_parallel_sets_labels()` from the `r CRANpkg("ggforce")` package but with the ability to
nudge labels at a fixed distance. It is especially useful when the labels are too long to fit
inside the bars depicting the discrete categories. A pull request of the nudge enhancement has
been submitted to its GitHub repository, `r Githubpkg("thomasp85/ggforce")`, awaiting approval.

```{r nudge-labels, fig.wide = TRUE, fig.asp = 0.8}
library(ggforce)
data <- as.data.frame(Titanic)
data <- gather_set_data(data, 1:4)

# Use nudge_x to offset and hjust = 0 to left-justify label
ggplot(data, aes(x, id = id, split = y, value = Freq)) +
  geom_parallel_sets(aes(fill = Sex), alpha = 0.3, axis.width = 0.1) +
  geom_parallel_sets_axes(axis.width = 0.1) +
  geom_parallel_sets_labs(colour = "red", size = 6, angle = 0,
                          nudge_x = 0.1, hjust = 0) +
  theme_bw(20)
```

## Create a 2-variable parallel sets diagram

The `plotParallel()` function uses the `r CRANpkg("ggforce")` package to produce a parallel
sets diagram for visualising interaction between 2 variables. The inputs are two character
vectors containing membership information.

The example below uses the `Titanic` dataset to show the class and age of the passengers.

```{r plot-parallel-1, fig.wide = TRUE, fig.asp = 0.4}
data <- as.data.frame(Titanic)
plotParallel(data$Class, data$Age, labels = c("class", "age"))
```

We can also use `plotParallel()` to show cell-specific features of a single-cell dataset, such 
as clustering and cell type assignment.

```{r plot-parallel-2, fig.wide = TRUE}
data(sce)

plotParallel(sce$label, sce$CellType, labels = c("Cluster", "Cell Type"),
             add_counts = TRUE, text_size = 4)
```

# Session information {-}

```{r}
sessionInfo()
```
