---
title: "Introduction to scRUtils"
author: "I-Hsuan Lin"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
package: scRUtils
output:
  BiocStyle::html_document:
    toc_float: yes
vignette: >
  %\VignetteIndexEntry{1. Introduction to scRUtils}
  %\VignetteEngine{knitr::rmarkdown}
  %VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(error = FALSE, warning = FALSE, message = TRUE, tidy = FALSE, 
		      cache = FALSE, dpi = 100)
library(BiocStyle)
library(DESeq2)
```

# Overview

`scRUtils` provides various utilities for visualising and functional analysis of RNA-seq data,
particularly single-cell dataset. It evolved from a collection of helper functions that were
used in our in-house scRNA-seq processing workflow.

The documentation of this package is divided into 5 sections:

1. [Introduction to scRUtils](https://ycl6.github.io/scRUtils/scRUtils-intro.html)
2. [General data visualisations](https://ycl6.github.io/scRUtils/basics.html)
3. [Single-cell visualisations](https://ycl6.github.io/scRUtils/singleCell.html)
4. [Markers and DEGs](https://ycl6.github.io/scRUtils/markers.html)
5. [Demo datasets](https://ycl6.github.io/scRUtils/demoData.html)

This vignette (#1) will introduce the package and the included demo datasets.

# Installation

The package is current available only on GitHub.

```{r, eval = FALSE}
devtools::install_github("ycl6/scRUtils")
```

# Load packages

To use `scRUtils` in a R session, load it using the `library()` command.

```{r load-packages, message = FALSE}
library(scRUtils)
```

# Use demo datasets

Demo datasets were created to demonstrate functions available in this package. The datasets can 
be loaded by using `data()`. Details on how the datasets were produced can be found in the 
following vignette: [Creating demo datasets](demoData.html).

## `cyclone` result

`phases_assignments` is a `list` object of 3, containing `cyclone` results performed on a 
simulated scRNA-seq dataset.

```{r load-cyclone}
data(phases_assignments)
str(phases_assignments)

table(phases_assignments[["phases"]])
```

## `findDoubletClusters` result

`dbl_results` is a `DFrame` object containing`findDoubletClusters` results performed on a 
simulated scRNA-seq dataset.

```{r load-dbl}
data(dbl_results)
dbl_results
```

## Processed simulated single-cell RNA-seq dataset

`sce` is a `SingleCellExperiment` object that has been processed with additional information 
stored.

```{r load-sce}
data(sce)
sce

# rowData
rowData(sce)

# colData
colData(sce)

# metadata
names(metadata(sce))
```

## edgeR's `TopTags` object

`res_edger` is a `TopTags` object containing differential expression analysis results of the 
`pasilla` dataset.

```{r load-res_edger}
data(res_edger)
DataFrame(res_edger)
```

## DESeq2's `DESeqResults` object

`res_deseq2` is a  `DESeqResults` object containing differential expression analysis results of 
the `pasilla` dataset.

```{r load-res_deseq2}
data(res_deseq2)
DataFrame(res_deseq2)
```

# Session information {-}

```{r}
sessionInfo()
```
